How to Not write Virtual Machines for Dynamic Languages

### 2007-05-27 talk/dyla2007/dyla.tex
Detail.
add a citation of the psyco home page too, differentiate better between citations of the psyco paper and psyco-the-program

These limitations of a C-based implementation often lead to more implementations
of the language. For example in addition to the main C implementation of Python
(``CPython'' \cite{cpy251}) there are a number of additional implementations
that want to fix some of the limitations of CPython. One of them is ``Stackless
Python'' \cite{stackless} which is a fork of CPython that adds micro-threading
capabilities to Python. One of the reasons for not incorporating it back into
CPython was that it was felt that they would make the implementation too
complex. Another implementation of the Python language is the Psyco \cite{Psyco}
project which adds a JIT-compiler to CPython. As all other re-implementations,
Psyco contains the Python semantics and needs to be kept synchronized with
CPython manually.

@@

An example where this mapping does not work too well is Prolog. While there
exist several implementations of Prolog on top of the JVM \cite{prologcafe}
\cite{DBLP:conf/jelia/Calejo04} and also one on .NET \cite{psharp},
they are not particular efficient, especially when compared to good Prolog VMs
in written in C. This is mostly because the Prolog execution model, which
involves backtracking and deep recursion does not fit the JVM and .NET very
well. Therefore the Prolog implementations on top of OO VMs resort to models
that is quite unnatural both for the OO VM and for Prolog.
\end{itemize}

@@

By far the fastest Python implementation, Psyco \cite{Psyco} contains a
hand-written language-specific dynamic compiler.  PyPy's translation
tool-chain is able to extend the generated VMs with an automatically
generated dynamic compiler similar to Psyco, derived from the
interpreter.  This is achieved by a pragmatic application of partial
evaluation techniques guided by a few hints added to the source of the
interpreter.  In other words, it is possible to produce a reasonably
good language-specific JIT compiler and insert it into a VM, alongside
with the necessary support code and the rest of the regular interpreter.

This result was one of the major goals and motivations for the whole
approach.  By construction, any code written in the dynamic language
runs correctly under the JIT.  Some very simple Python examples run more
than 100 times faster.  At the time of this writing this is still rather
experimental, and the techniques involved are well beyond the scope of
the present paper.  The reader is referred to \cite{D08.2} for more
information.

\section{Related Work}
XXX

>>>>

These limitations of a C-based implementation often lead to more implementations
of the language. For example in addition to the main C implementation of Python
(``CPython'' \cite{cpy251}) there are a number of additional implementations
that want to fix some of the limitations of CPython. One of them is ``Stackless
Python'' \cite{stackless} which is a fork of CPython that adds micro-threading
capabilities to Python. One of the reasons for not incorporating it back into
CPython was that it was felt that they would make the implementation too
complex. Another implementation of the Python language is the Psyco
\cite{psyco-software}
project which adds a JIT-compiler to CPython. As all other re-implementations,
Psyco contains the Python semantics and needs to be kept synchronized with
CPython manually.

@@

An example where this mapping does not work too well is Prolog. While there
exist several implementations of Prolog on top of the JVM \cite{prologcafe}
\cite{InterProlog} and also one on .NET \cite{psharp},
they are not particular efficient, especially when compared to good Prolog VMs
in written in C. This is mostly because the Prolog execution model, which
involves backtracking and deep recursion does not fit the JVM and .NET very
well. Therefore the Prolog implementations on top of OO VMs resort to models
that is quite unnatural both for the OO VM and for Prolog.
\end{itemize}

@@

By far the fastest Python implementation, Psyco \cite{psyco-software} contains a
hand-written language-specific dynamic compiler.  PyPy's translation
tool-chain is able to extend the generated VMs with an automatically
generated dynamic compiler that uses techniques similar to those of Psyco
\cite{Psyco-paper}, derived from the
interpreter.  This is achieved by a pragmatic application of partial
evaluation techniques guided by a few hints added to the source of the
interpreter.  In other words, it is possible to produce a reasonably
good language-specific JIT compiler and insert it into a VM, alongside
with the necessary support code and the rest of the regular interpreter.

This result was one of the major goals and motivations for the whole
approach.  By construction, the JIT does not get out of sync when the
language evolves, and any code written in the dynamic language runs
correctly under the JIT.  Some very simple Python examples run more than
100 times faster.  At the time of this writing this is still rather
experimental, and the techniques involved are well beyond the scope of
the present paper.  The reader is referred to \cite{D08.2} for more
information.
